\textsc{RILS}-\textsc{ROLS} : Robust Symbolic Regression via Iterated Local Search and Ordinary Least Squares

### 2022-11-28 paper_resources/doc/paper.tex
Section 5: polishing

\begin{center}
	\begin{tikzpicture}
		\begin{axis}[
			xlabel=\textbf{model size},
			ylabel=\textbf{$R^2 > 0.999$ percentage},
			xmin=0, xmax=4,
			ymin=20, ymax=100,
			xtick={1,2,3},
			xticklabels={small,medium,large},   % <---
			ytick={20,30,...,100}
			]
			\addplot[smooth,mark=*,blue] plot coordinates {
				(1,98.33)
				(2,62.5)
				(3,58.95)
			};
			\addlegendentry{No noise}

			\addplot[smooth,color=red,mark=x]
			plot coordinates {
				(1,93.33)
				(2,53.75)
				(3,40)
			};
			\addlegendentry{Noise level 0.001}

			\addplot[smooth,color=green,mark=o]
			plot coordinates {
				(1,65)
				(2,31.25)
				(3,20)
			};
			\addlegendentry{Noise level 0.01}
		\end{axis}
	\end{tikzpicture}
	\captionof{figure}{Percentages of solutions having $R^2 > 0.999$ for varying levels of noise and formulae sizes}
	\label{fig:compR2_noise_size}
\end{center}

@@

In this paper, we dealt with solving the well--known Symbolic regression (SR)  problem. SR problem is a generalization of more interdisciplinary known problems of linear and logistic regression.    
We proposed a meta-heuristic approach, called \textsc{RILS}-\textsc{ROLS}. Our approach is built upon the Iterated local search (ILS) scheme. The first crucial concept within this scheme is utilizing the ordinary least square method to efficiently determine internal coefficients that are the best fitting to the candidate solution. Thus, the search process is primarily focused on considering already promising solutions in the continuous part of the search space rather than applying enhanced heuristics to search for those coefficients. Another key aspect is the utilization of a carefully constructed fitness function in the search, which appeared as a combination of three important scores: RMSE, $R^2$  scores and the size of the model (solution).  Last but not least, a carefully constructed local search is applied systematically exploring solutions obtained as 1--perturbations of the considered solution.  1--perturbations of a solution are based on applying the six per-node change operations on the solution's tree representation.

\textsc{RILS-ROLS} method is compared to the 14 other competitor algorithms from the literature on the two ground truth benchmark sets from literature, namely \textsc{Feynman} and \textsc{Strogatz}.  Additionally, we introduced a non-biased benchmark set of randomly generated instances, labelled by \textsc{Random}.  The experimental evaluation confirmed the efficiency of our method which produced the best average ranking results in terms of exact percentage when compared to all other competitors on the two ground-truth benchmarks in the presence of all three different levels of noise in the input data. In all cases, new best exact percentages are obtained using \textsc{RILS-ROLS} method.   More in details,
our algorithm was able to obtain the right model on 60.62\%, 42.08\%, and 34.77\% problem instances in the presence of the noise level of 0.0, 0.001, and 0.01, respectively. Just for the shake of comparisons, the second best approach \textsc{AI-Feynman}, when the noise level of 0.0 and 0.001 is included, obtained an exact percentage rate of 52.65\%, and 31.89\%, respectively. In case of the highest noise level, the second best approach was \textsc{Afp-Fe}, obtaining an exact percentage rate of 20\%. Additionally, the scalability and robustness of this method are checked and confirmed on the unbiased \textsc{Random} benchmark set in terms of a few different criteria such as the size of the model, the number of variables included in exact models, and obtained average runtimes.

In future work, one could think of constructing a hybrid of \textsc{RILS-ROLS} method with some other meta-heuristics to further boost the quality of the obtained results. For example, replacing ILS scheme of \textsc{RILS-RILS} with a more general Variable neighbourhood search (VNS) scheme is a reasonable option to give it a try. VNS   utilizes a more stronger and systematic system of solutions diversification and thus escaping from local optima could be much easier and the convergence much faster.  Additionally, as proof of the efficiency and benefits of our algorithm, this tool could be applied to solving problems from various problem domains out of the area of computer science, for example from physics or chemistry, to help for setting up reasonable hypotheses or to obtain valuable insights on experimental evaluations.

>>>>

\begin{center}
	\begin{tikzpicture}
		\begin{axis}[
			xlabel=\textbf{model size},
			ylabel=\textbf{$R^2 > 0.999$ percentage},
			xmin=0, xmax=4,
			ymin=20, ymax=100,
			xtick={1,2,3},
			xticklabels={small,medium,large},   % <---
			ytick={20,30,...,100}
			]
			\addplot[smooth,mark=*,blue] plot coordinates {
				(1,98.33)
				(2,62.5)
				(3,58.95)
			};
			\addlegendentry{$\alpha=0.00$}

			\addplot[smooth,color=red,mark=x]
			plot coordinates {
				(1,93.33)
				(2,53.75)
				(3,40)
			};
			\addlegendentry{$\alpha=0.001$}

			\addplot[smooth,color=green,mark=o]
			plot coordinates {
				(1,65)
				(2,31.25)
				(3,20)
			};
			\addlegendentry{$\alpha= 0.01$}
		\end{axis}
	\end{tikzpicture}
	\captionof{figure}{Percentages of solutions having $R^2 > 0.999$ for varying levels of noise and formulae sizes}
	\label{fig:compR2_noise_size}
\end{center}

@@

In this paper, we dealt with solving the well--known Symbolic regression (SR)  problem. SR problem is a generalization of more interdisciplinary known problems such as linear and logistic regression.    
We proposed a meta-heuristic approach, called \textsc{RILS}-\textsc{ROLS}, which is built upon the Iterated local search (ILS) scheme. The first crucial concept within this scheme is utilizing the ordinary least square method to efficiently determine best-fitting internal coefficients to the candidate solution. % Thus, the search process is primarily focused on considering already promising solutions in the continuous part of the search space rather than applying enhanced heuristics to search for those coefficients. 
Another key aspect is the utilization of a carefully constructed fitness function in the search, which appeared as a combination of three important scores: RMSE, $R^2$  scores and the size of the model (solution).  Last but not least, a carefully constructed local search is applied systematically exploring solutions obtained as 1--perturbations of the considered solution, represented as trees.  1--perturbations of a solution are based on applying the six per-node change operations on the solution's tree representation.

\textsc{RILS-ROLS} method is compared to the 14 other competitor algorithms from the literature on the two ground--truth benchmark sets from the literature, namely \textsc{Feynman} and \textsc{Strogatz}.  Additionally, we introduced a non-biased benchmark set of randomly generated instances, labelled by \textsc{Random}.  The experimental evaluation confirmed the efficiency of our method which produced the best average ranking results in terms of exact percentage when compared to all other competitors on the two ground-truth benchmarks in the presence of all three different levels of noise in the input data. In all cases, new best exact percentages are obtained by our \textsc{RILS-ROLS} method.   More in details,
our algorithm was able to obtain the right model on 60.62\%, 42.08\%, and 34.77\% runs over all considered problem instances in the presence of different noise levels: 0.0, 0.001, and 0.01, respectively. For the shake of comparisons, the second best approach \textsc{AI-Feynman}, when the noise level of 0.0 and 0.001 is included, obtains the exact percentages of 52.65\%, and 31.89\%, respectively. In case of the highest noise level, the second best approach was \textsc{Afp-Fe}, obtaining an exact percentage rate of 20\%. Additionally, the scalability and robustness of \textsc{RILS}-\textsc{ROLS} method are checked and confirmed on the unbiased \textsc{Random} benchmark set in terms of a few different criteria such as the size of the model, the number of variables included in exact models, and obtained average runtimes.

In future work, one could think of constructing a hybrid of \textsc{RILS-ROLS} method with some other meta-heuristics to further boost the quality of the obtained results. For example, replacing ILS scheme of \textsc{RILS-RILS} with a more general Variable neighbourhood search (VNS) scheme is a reasonable option to give it a try. VNS   utilizes a more stronger and systematic system of solutions diversification and thus escaping from local optima could be much easier and the convergence much faster.  Additionally, as proof of the efficiency and usefulness of our algorithm, this tool could be applied to solving problems from various problem domains out of the area of computer science, for example from physics or chemistry, to help for setting up reasonable hypotheses or to obtain valuable insights on obtained experimental evaluations.
